---
title: "Processing raw maple data"
author: "Erika Barthelmess"
date: "8/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
During summmer 2019, we built out a new data entry portal on the Nature Up North website to allow easier submission of maple data.  We would like to process the historical data into a csv that can be easily batch uploaded to the website so that folks have access to all of the data.  That's what this code is for.

## Set up the workspace

```{r}
rm(list = ls())
library(ggplot2)
library(tidyverse)
```

## Now pull in the data
I am taking data from the csv file called "all_maple_monitoring_through_fall2017" which is found in Nature Up North/Citizen Science/MapleMonitoring/Data on dropbox.

```{r}
Trees <- read.csv("/Volumes/External Drive/Dropbox/R/NUN/Maple Data/TreesProcessed.csv", stringsAsFactors=FALSE)
# library(readr)
# Trees <- read_csv("all_maple_monitoring_through_fall2017.csv", 
#     col_types = cols(Circumference = col_number(), 
#         Date = col_date(format = "%m/%d/%y"), 
#         Julian =col_number(),
#         Flowering = col_character(), Latitude = col_number(), 
#         LeafBud = col_character(), LeafUnfurled = col_character(), 
#         Longitude = col_number(), Observer = col_character(), 
#         `Tree ID` = col_character(), Username = col_character()))
# 
# #for some reason when I just did a straight read_csv, the columns indicated in the col_types section didn't get pulled in with the rest of the data...this seems to fix it.  Or use read.csv from base seems fine...
```
There are 2108 observations of 24 variables.

## Now do some data cleanup.
###Get rid of the extra date column
```{r}
Trees$X<-NULL
```
### Make sure that levels of factor variables are correct.

Create factor variables
```{r}
vars<-c("Season", "Species", "Urbanization", "Habitat", "Shading", "LeafColor", "LeafDropping", "Fruit", "LeafBud", "LeafUnfurled", "Flowering")

Trees[vars]<-lapply(Trees[vars],as.factor)
```
###Now check for levels of each factor 
Make sure there are not multiple spellings of factor levels, etc. Repair any problems as they arise.

```{r}
levels(Trees$Season)
```
Season looks good.
```{r}
levels(Trees$Species)
```

Species also looks good.
```{r}
levels(Trees$Urbanization)
```
Looks good.
```{r}
levels(Trees$Habitat)
```
Habitat also looks good.
```{r}
levels(Trees$Shading)
```

```{r}
levels(Trees$LeafColor)
```
```{r}
levels(Trees$LeafDropping)
```
```{r}
levels(Trees$Fruit)
```
```{r}
levels(Trees$LeafBud)
```
```{r}
levels(Trees$LeafUnfurled)
```
```{r}
levels(Trees$Flowering)
```


Looks good.  None of the factor variables have incorrect levels assigned.

###Now check for errors in numeric variables

####Examine latitude and longitude
Latitude: First look at the ranges of values
```{r}
summary(Trees$Latitude)
```
All the latitude values appear to be ok, though some may be wrong.  The -76 is probabaly a lattitude.  But I'm not going to change it.


Longitude: First look at the ranges of values

```{r}
summary(Trees$Longitude)
```
Not all longitude values are negative. Need to fix that. Some are unusual but we're not going to change any data at this point.

Make them all negative

```{r}
Trees$Longitude<-abs(Trees$Longitude)
Trees$Longitude<-Trees$Longitude*-1
summary(Trees$Longitude)
```


####Other numeric variables
Can now move on to next numeric variable, circumference.
Look at range of tree circumferences.
```{r}
summary(Trees$Circumference)
```
Let's make a plot so see how many really big numbers there are (e.g. 241.00!)
```{r}
ggplot(Trees, aes(Circumference))+
  geom_histogram()
```
Ok.  Numeric variables look good.  


###Now populate other columns
####Serial
Add a unique ID to each row of data
```{r}
# Trees$Serial<-(1:2108)
```
####Julian
Now see if I can add a julian date.  Note that some of the dates in the date fields are March, rather than a fall date.  I think these are the dates the data were contributed rather than the dates the observations were made.  I am not changing them.  The data are the data.

And now get the Julian date (you'll need the lubridate package if you don't bring them in as dates)
```{r}
library(lubridate)
Trees$Date<-ymd(Trees$Date)
Trees$Julian<-yday(Trees$Date)
```
## Final Steps - Concatenate and export file
We need a unique ID per tree, but we don't want each observation of the same tree to count as a different tree.  Thus, we are concatenating a bunch of variables into a string that should be unique to each tree, but not each observation, so that we can identify all of the observations that appear to be from the same tree.

To do so, we will create two new columns.  First we will create a LongLat column that is the longitude value minus the latitude value.  Then we will concatenate the columns LongLat, Tree ID, Species, Latitude, Longitude, Urbanization, Habitat, Shading and Circumference.  These will be used to determine which rows are unique trees.

```{r}
Trees$LongLat<-Trees$Longitude - Trees$Latitude
Trees$UniqueID<-paste(Trees$LongLat,Trees$Observer,Trees$`Tree ID`, Trees$Species, Trees$Latitude, Trees$Longitude, Trees$Urbanization, Trees$Habitat, Trees$Shading, Trees$Circumference, sep = "_")
```
Now write the file to a new csv
```{r}
write.csv(Trees, "TreesProcessed.csv")
```





