---
title: "Processing raw maple data"
author: "Erika Barthelmess"
date: "8/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
During summmer 2019, we built out a new data entry portal on the Nature Up North website to allow easier submission of maple data.  We would like to process the historical data into a csv that can be easily batch uploaded to the website so that folks have access to all of the data.  That's what this code is for.

## Set up the workspace

```{r}
rm(list = ls())
library(ggplot2)
library(tidyverse)
```

## Now pull in the data
I am taking data from the excel file called "all_maple_monitoring_through_fall2017.xlsx" which is found in Nature Up North/Citizen Science/MapleMonitoring/Data on dropbox.

```{r}
library(readr)
Trees <- read_csv("FallTrees.csv")
```
There are 1502 observations of 24 variables.

## Now do some data cleanup.

### Begin by getting rid of the spring data observations
```{r}
Trees<-select(Trees, -starts_with("Leaf"))
Trees<-select(Trees, -Fruit)
```

### Continue by making sure that levels of factor variables are correct.

Create factor variables
```{r}
Trees$Season<-as.factor(Trees$Season)
Trees$Species<-as.factor(Trees$Species)
Trees$Urbanization<-as.factor(Trees$Urbanization)
Trees$Habitat<-as.factor(Trees$Habitat)
Trees$Shading<-as.factor(Trees$Shading)
```
###Now check for levels of each factor 
Make sure there are not multiple spellings of factor levels, etc. Repair any problems as they arise.

```{r}
levels(Trees$Season)
```
Season looks good.
```{r}
levels(Trees$Species)
```

Species also looks good.
```{r}
levels(Trees$Urbanization)
```
Looks good.
```{r}
levels(Trees$Habitat)
```
Habitat also looks good.
```{r}
levels(Trees$Shading)
```
Looks good.  None of the factor variables have incorrect levels assigned.



###Now check for errors in numeric variables
####Examine latitude and longitude
Latitude: First look at the ranges of values
```{r}
summary(Trees$Latitude)
```
All the latitude values appear to be ok, though some may be wrong.


Longitude: First look at the ranges of values

```{r}
summary(Trees$Longitude)
```
All longitude values are negative.  Some are unusual, (e.g. -90, -7) but we're not going to change any data at this point.
####